Paper Analyses

### 2022-07-11 DataSetup.Rmd
Cleaned up data setup code and added new variables, added code for calculating difference scores

---
title: "DataSetup"
output: html_document
---

@@

Load Structural Data
```{r}
#Load list of included subjects
struc_include<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:12) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include),
        imgincl_t2w_include = as.numeric(imgincl_t2w_include)) %>% 
  filter(imgincl_t1w_include == 1)

#Load structural data
#Filter for volume & area measurements
smri_1 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area"))

smri_baseline <- smri_1%>% 
  filter(eventname=="baseline_year_1_arm_1")

smri_followup <- smri_1%>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#T1 intensity, might not need
smri_2 <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip20201.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include$src_subject_id)

#Rename data to fit naming conventions
#????

@@

Load Other Data
```{r}
#Puberty
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title")

youth_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphy01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")

parent_pub_sum <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ssphp01.txt") %>%
  filter(!collection_title=="collection_title")

saliva <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_hsss01.txt") %>%
  filter(!collection_title=="collection_title")

#parent sum scores abcd_mhp01
#youth sum scores abcd_mhy01

#Cognition
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, nihtbx_fluidcomp_fc, nihtbx_cryst_fc, nihtbx_totalcomp_fc)

#Height/weight metrics
physical <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ant01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, anthroheightcalc, anthroweightcalc)

#Sexual attraction and behavior
gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")

#Grade level, parent report
grade <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(subjectkey, src_subject_id, interview_date, interview_age, sex, eventname, demo_ed_v2)
```

Template for renaming sMRI variables
```{r, eval=F}

@@

# no cerebellum cort vol?

@@

#renamed_smri <- subset(renamed_smri, select = -c(269,279))

#Convert to numeric
renamed_smri_baseline[,6:279] <- sapply(renamed_smri_baseline[,6:279],as.numeric)
renamed_smri_baseline <- renamed_smri_baseline %>% 
  mutate(interview_age = as.numeric(interview_age)/12)

@@

Feature list template
```{r, eval=F}
#Template
FS_InterCranial_Vol = smri_vol_scs_intracranialv, #technically inter vs intra
FS_BrainSeg_Vol
FS_BrainSeg_Vol_No_Vent
FS_BrainSeg_Vol_No_Vent_Surf

@@

FS_Total_GM_Vol #maybe just calculate?

@@

FS_SupraTentorial_Vol_No_Vent  #smri_vol_scs_allventricles is all ventricles, subtract?
FS_SupraTentorial_No_Vent_Voxel_Count
FS_Mask_Vol
FS_BrainSegVol_eTIV_Ratio
FS_MaskVol_eTIV_Ratio

@@

FS_L_Vessel_Vol
FS_L_ChoroidPlexus_Vol

@@

FS_R_Vessel_Vol
FS_R_ChoroidPlexus_Vol
FS_OpticChiasm_Vol

@@

@@

>>>>

---
title: "FYP Redo"
author: "Lucy Whitmore"
date: "5/19/2022"
output: html_document
---

@@

Load Structural Data
```{r}
#Load list of included subjects
struc_include_baseline<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="baseline_year_1_arm_1")

struc_include_followup<-rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_imgincl01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(1:11) %>% 
  mutate(interview_age = as.numeric(interview_age),
        imgincl_t1w_include = as.numeric(imgincl_t1w_include)) %>% 
  filter(imgincl_t1w_include == 1) %>% 
  filter(eventname=="2_year_follow_up_y_arm_1")

#Load structural data
#Filter for volume & area measurements
smri_baseline <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="baseline_year_1_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_baseline$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

smri_followup <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_smrip10201.txt") %>%
  filter(eventname=="2_year_follow_up_y_arm_1") %>% 
  filter(!collection_title=="collection_title") %>% 
  filter(src_subject_id%in%struc_include_followup$src_subject_id) %>% 
  select(subjectkey, src_subject_id, interview_age, sex, eventname, matches("vol|area")) %>% 
  select(-contains(c('cf'))) %>% #remove genetically derived parcellations
  select(-c(smri_area_cdk_totallh, smri_area_cdk_totalrh, smri_area_cdk_total, smri_vol_scs_lesionlh, smri_vol_scs_lesionrh, smri_vol_scs_wmhint, smri_vol_scs_wmhintlh, smri_vol_scs_wmhintrh, smri_vol_scs_wholeb, smri_vol_scs_latventricles, smri_vol_scs_allventricles, smri_vol_scs_cbwmatterrh, smri_vol_scs_cbwmatterlh))

#renamed_smri_baseline<-rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/FYP/renamed_smri_baseline.Rda")

@@

Rename Columns Baseline
```{r}

@@

FS_R_Cerebellum_Cort_Vol = smri_vol_scs_crbcortexrh,

@@

#Convert to numeric
renamed_smri_baseline[,6:ncol(renamed_smri_baseline)] <- sapply(renamed_smri_baseline[,6:ncol(renamed_smri_baseline)],as.numeric)
renamed_smri_baseline <- renamed_smri_baseline %>% 
  mutate(interview_age = as.numeric(interview_age)/12)

@@

smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180] 
```

Rename Columns Follow-up
```{r}
renamed_smri_followup <- smri_followup %>% 

@@

@@

@@

@@

@@

#Create empty columns to match feature list
renamed_smri_followup[,"FS_BrainSeg_Vol"] <- NA
renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent"] <- NA
renamed_smri_followup[,"FS_BrainSeg_Vol_No_Vent_Surf"] <- NA
renamed_smri_followup[,"FS_SupraTentorial_Vol_No_Vent"] <- NA
renamed_smri_followup[,"FS_SupraTentorial_No_Vent_Voxel_Count"] <- NA
renamed_smri_followup[,"FS_Mask_Vol"] <- NA
renamed_smri_followup[,"FS_BrainSegVol_eTIV_Ratio"] <- NA
renamed_smri_followup[,"FS_MaskVol_eTIV_Ratio"] <- NA
renamed_smri_followup[,"FS_L_Vessel_Vol"] <- NA
renamed_smri_followup[,"FS_L_ChoroidPlexus_Vol"] <- NA
renamed_smri_followup[,"FS_R_Vessel_Vol"] <- NA
renamed_smri_followup[,"FS_R_ChoroidPlexus_Vol"] <- NA
renamed_smri_followup[,"FS_OpticChiasm_Vol"] <- NA
renamed_smri_followup[,"FS_Total_GM_Vol"] <- NA

#Convert to numeric
renamed_smri_followup[,6:ncol(renamed_smri_followup)] <- sapply(renamed_smri_followup[,6:ncol(renamed_smri_followup)],as.numeric)
renamed_smri_followup <- renamed_smri_followup %>% 
  mutate(interview_age = as.numeric(interview_age)/12)

save(renamed_smri_followup, file="renamed_smri_followup.Rda")

smri_abcd_model_followup <- renamed_smri_followup[, 1:180] 
```

Load Other Variables
```{r}
#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

#Raw scores, parent
parent_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ppdms01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(5:27) %>% 
  select(-c(interview_date, pds_select_language___1)) %>% 
  mutate_at(c(5:21), as.numeric)

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_rawscore, nihtbx_totalcomp_cs, nihtbx_totalcomp_itmcnt) %>% 
  mutate_at(c(2, 4:9), as.numeric)

#nihtbx_totalcomp_rawscore total cognition raw score
#nihtbx_totalcomp_cs total cognition computed score
#	nihtbx_totalcomp_itmcnt cognition total composite item count
# might need to adjust for tasks completed?

social <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(-c(subjectkey, interview_date, interview_age, gish2_device)) %>% 
  mutate_at(c(4:16), as.numeric)

#calculate gish score
```

## Format Puberty Data 
```{r}
#two or more missing values = exclude
#PDS mean score is a continuous value calculated by averaging across PDS items following Herting et al. (2021). Items on the PDS were considered missing if participants answered, “I don’t know” or “refuse to answer”, if the response was left blank, or if the response value was outside of the expected range (1 to 4). Participants with two or more missing answers were excluded from the PDS score calculation (following recommendations from (Herting et al., 2021)

### Youth-report, removes everyone w/o complete data
youth_pds[youth_pds == 777] <- NA
youth_pds[youth_pds == 999] <- NA

# Multiwave
youth_pds_multiwave <- youth_pds %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)  %>% 
  select(src_subject_id, sex, youth_sum, youth_mean)

youth_pds_baseline <- youth_pds_multiwave %>% 
  filter(eventname == "baseline_year_1_arm_1")

youth_pds_followup <- youth_pds_multiwave %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")

#### Parent-report, removes everyone w/o complete data
parent_pds[parent_pds == 777] <- NA
parent_pds[parent_pds == 999] <- NA

#Baseline

parent_pds_multiwave <- parent_pds %>% 
    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5) %>% 
  select(src_subject_id, sex, parent_sum, parent_mean)

parent_pds_baseline <- parent_pds_multiwave %>% 
    filter(eventname == "baseline_year_1_arm_1")

parent_pds_followup <- parent_pds_multiwave%>% 
    filter(eventname == "2_year_follow_up_y_arm_1")

#### Cognition

#Baseline
cognition_baseline <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1")

#Follow-Up
cognition_followup <- cognition %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")

#Multiwave
cognition_multiwave <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1")
```

Add age to brainAGE estimates
```{r}
corrected_baseline_brainage <- brain_age_corrected_df_baseline
corrected_followup_brainage <- brain_age_corrected_df_followup

corrected_baseline_brainage$src_subject_id <- smri_baseline$src_subject_id

corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id

@@

Split Samples for Analysis
```{r}

brainage_pds_baseline <- corrected_baseline_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_baseline) %>% 
  left_join(parent_pds_baseline) %>% 
  left_join(cognition_baseline)

#Split sample
smp_size_baseline <- floor(0.5 * nrow(brainage_pds_baseline))

## set the seed to make your partition reproducible
set.seed(123)
sample_ind_baseline <- sample(seq_len(nrow(brainage_pds_baseline)), size = smp_size_baseline)

sample_1_baseline <- brainage_pds_baseline[sample_ind_baseline, ]
sample_2_baseline <- brainage_pds_baseline[-sample_ind_baseline, ]

#t test for age differences
t.test(sample_1_baseline$interview_age, sample_2_baseline$interview_age)

#Save samples
save(brainage_pds_baseline, file="BrainagePDSBaseline.Rda")
save(sample_1_baseline, file="Sample1Baseline.Rda")
save(sample_2_baseline, file = "Sample2Baseline.Rda")

# Follow-Up
brainage_pds_followup <- corrected_followup_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_followup) %>% 
  left_join(parent_pds_followup) %>% 
  left_join(cognition_followup)

#Split sample
smp_size_followup <- floor(0.5 * nrow(brainage_pds_followup))

## set the seed to make your partition reproducible
set.seed(123)
sample_ind_followup <- sample(seq_len(nrow(brainage_pds_followup)), size = smp_size_followup)

sample_1_followup <- brainage_pds_followup[sample_ind_followup, ]
sample_2_followup <- brainage_pds_followup[-sample_ind_followup, ]

#t test for age differences
t.test(sample_1_followup$interview_age, sample_2_followup$interview_age)

#Save samples
save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
save(sample_1_followup, file="Sample1FollowUp.Rda")
save(sample_2_followup, file = "Sample2FollowUp.Rda")
```

#################
Make Dataframe for ABCD Model Training & Testing
################
```{r}

#smri_abcd_model_baseline <- renamed_smri_baseline[, 1:180]

training_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_1_baseline$src_subject_id) %>% 
  select(-c(subjectkey, src_subject_id, sex, eventname))

analysis_sample_baseline <- filter(smri_abcd_model_baseline, src_subject_id %in% sample_2_baseline$src_subject_id)

save(training_sample_baseline, file="training_sample_baseline.Rds")
save(analysis_sample_baseline, file="analysis_sample_baseline.Rds")

#################
Make Dataframe of sMRI features for combined wave model
################
```{r}

#Pull 1800 total
#900 from each
#Pull 900 from baseline
#filter those 900 out of the followup and then pull 900

## set the seed to make your partition reproducible
set.seed(42)
#Pull 900 from baseline
sample_ind_abcd_baseline <- sample(seq_len(nrow(smri_abcd_model_baseline)), size = 900)

abcd_train_baseline <- smri_abcd_model_baseline[sample_ind_abcd_baseline, ]
abcd_test_baseline <- smri_abcd_model_baseline[-sample_ind_abcd_baseline, ]

#filter baseline participants out of follow-up
smri_abcd_model_followup_filtered <- smri_abcd_model_followup %>% 
  filter(!src_subject_id %in% abcd_train_baseline$src_subject_id)

set.seed(42)
#Pull 900 from followup
sample_ind_abcd_followup <- sample(seq_len(nrow(smri_abcd_model_followup_filtered )), size = 900)

abcd_train_followup <- smri_abcd_model_followup[sample_ind_abcd_followup, ]
abcd_test_followup <- smri_abcd_model_followup[-sample_ind_abcd_followup, ]

abcd_train_multiwave <- rbind(abcd_train_baseline, abcd_train_followup) %>% 
  select(-c(subjectkey, src_subject_id, sex, eventname))
save(abcd_train_multiwave, file="abcd_train_multiwave.Rds")

abcd_test_multiwave <- rbind(abcd_test_baseline, abcd_test_followup)
save(abcd_test_multiwave, file="abcd_test_multiwave.Rds")

##############
Add Other Variable to Multiwave Test
##############
Add age to brainAGE estimates
```{r}
#load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDMultiwaveBrainAgeCorrected.Rda")

corrected_multiwave_brainage <- brain_age_corrected_df_multiwave

corrected_multiwave_brainage$src_subject_id <- abcd_test_multiwave$src_subject_id

#combine waves of youth pds

#combine waves of parent pds

#combine waves of cognition

#change
brainage_pds_multiwave <- corrected_multiwave_brainage[, c(6,1:5)] %>% 
  left_join(youth_pds_multiwave) %>% 
  left_join(parent_pds_multiwave) %>% 
  left_join(cognition_multiwave)

### 2022-07-11 PaperAnalyses&Figures.Rmd
Cleaned up data setup code and added new variables, added code for calculating difference scores

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_1_baseline)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean + interview_age, data = sample_2_baseline)

@@

@@

@@

summary(baseline_cognition_1_m)
summary(baseline_cognition__m2)

>>>>

# brainage ~ youth mean
baseline_puberty_y_mean_1 <- lm(corrected_gap ~ youth_mean, data = sample_1_baseline)
baseline_puberty_y_mean_2 <- lm(corrected_gap ~ youth_mean, data = sample_2_baseline)

@@

#baseline_puberty_y_mean_1_pred <- lm(corrected_pred ~ youth_mean, data = sample_1_baseline)
#baseline_puberty_y_mean_2_pred <- lm(corrected_pred ~ youth_mean, data = sample_2_baseline)

#baseline_puberty_y_mean_1_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_1_baseline_months)
#baseline_puberty_y_mean_2_m_pred <- lm(corrected_pred ~ youth_mean + interview_age_m, data = sample_2_baseline_months)

@@

#summary(baseline_puberty_y_mean_1_pred)
#summary(baseline_puberty_y_mean_2_pred)

@@

summary(baseline_cognition_1_m)
summary(baseline_cognition_2_m)

Basic Prediction Plots
```{r}

df <- ggpredict(baseline_puberty_y_mean_1_m, terms = c("youth_mean","interview_age_m"))
ggplot(df, aes(x, predicted)) + 
    geom_line(aes(linetype=group, color=group)) +
    geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill=group), alpha=0.15) +
    scale_linetype_manual(values = c("solid", "dashed", "dotted"))

plot_model(baseline_puberty_y_mean_1_m, type="pred", terms = "youth_mean", title = "Predicted Gap", colors = "#cc5968") +
  xlab("Youth-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_y_mean_2_m, type="pred", terms = "youth_mean", title = "Predicted Gap") +
  xlab("Youth-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_p_mean_1_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
  xlab("Parent-Report Puberty") +
  ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_puberty_p_mean_2_m, type="pred", terms = "parent_mean", title = "Predicted Gap") +
    xlab("Parent-Report Puberty") +
    ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_cognition_1_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
   xlab("Age-Corrected Cognition Score") +
    ylab("Predicted Gap") +
  theme_sjplot()

plot_model(baseline_cognition_2_m, type="pred", terms = "nihtbx_totalcomp_agecorrected", title = "Predicted Gap")+
    xlab("Age-Corrected Cognition Score") +
    ylab("Predicted Gap") +
  theme_sjplot()
